\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for Marginal Distributions of Stationary Time Series

### 2024-04-02 manuscript/manuscript.tex
few manuscript changes
manuscript tidying up

The Kolmogorov-Smirnov (KS) statistic is widely used to test if a sample is
from a given distribution. This study demonstrates how non-parametric 
block bootstrap can be 
used to approximate the KS statistic in situations where parameters are 
not specified and the data are serially dependent. We demonstrate the 
theoretical justication for this method, then through simulation we verify that
it holds it size under the null distribution and is powerful when the null 
hypothesis is false. Finally, we demonstrate applications of the method to
the S\&P 100 and S\&P 500 indices stock returns.

@@

The standard one-sample Kolmogorov-Smirnov (KS) test is widely
recoganized as effective good-of-fit test for continuous distributions.
Consider $X_1, ..., X_n$, a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ that $X_i$'s follow a specific
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. This distribution function is
precisely computable using contemporary statistical software
\citep{marsaglia2003evaluating}. The versatility of the KS test allows its
application across various domains, such as analyzing cosmic microwave
background radiation \citep{naess2012application}, monitoring the count rate of
radioactive data \citep{aslam2020introducing}, gear condition monitoring
\citep{andrade2001gear}, studying images of breast cancer tumors
\citep{demidenko2004kolmogorov}, and examining financial markets
\citep{lux2001turbulence}.

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} addresses this issue through the
application of basic non-parametric bootstrap methods.
Another commonly encountered situation is when the data are serially
dependent. The KS test is frequently misused without accounting for the
assumption of independence. In an investigation of performance variation in
high-performance computing systems, \citet{tuncer2019ieee} did not describe
how or if they accounted for serial dependence when applying the two-sample
KS test.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

@@

The remainder of this paper is structured as follows: 
Section~\ref{sec:methods} provides an overview of the block bootstrap 
procedure and introduces the bias correction methodology for the 
non-parametric block bootstrap KS test. Section~\ref{sec:simu} is divided 
into two parts; initially, we evaluate the KS test's ability to maintain 
its size, i.e., its consistency in not rejecting the null hypothesis when 
it is indeed true. Subsequently, we examine the test's power, assessing 
whether it can reject the null hypothesis when it is false. Practical 
applications of our method are presented in Section~\ref{sec:real}, where 
we apply the approach to analyze if S\&P 100 and S\&P 500 index stock return data adhere to either the Normal 
or the Student's $t$ distribution in Subsection~\ref{sec:sp100} and
\ref{sec:sp500}. The 
paper concludes with Section~\ref{sec:conclusion}, offering final thoughts 
and remarks.

@@

First, let us review how the KS statistic is typically computed for a sample
with no serial dependence and specified parameters.
Let $\hat\theta_n$ denote the fitted parameters for the hypothesized 
distribution fitted onto $X_t$, and let 
$F_n$ denote the empirical distribution function based on $X_1,...,X_n$.
Let
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the
goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

@@

In this simulation study, we aim to show that under the null
distribution, our test rejects the $H_0$ at the significance level, 
and that under the alternate distribution, our test rejects the $H_0$ at a high
rate. If
both of these things are true, we can state that the method works
reasonably well.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}$.
\mc{my explanation needs work}

@@

\subsection{SP500}
\label{sec:sp500}

Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices for the S\&P 500 index. We
observed how the method performs using increasingly smaller durations of
time.
We used the last 5 years of data: January 1st, 2019 to December 31st, 2023,
4 years of data: January 1st, 2020 to December 31st, 2023,
3 years of data: January 1st, 2021 to December 31st, 2023,
2 years of data: January 1st, 2022 to December 31st, 2023,
and 1 year of data: January 1st, 2023
to December 31st, 2023.
We approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
We used $B = 10,000$ bootstrap 
replicates in this analysis.
The p-values for these tests are highlighted in 
Tables~\ref{table:SP5001}--\ref{table:SP5005}.

\input{tables/sp500_pval_1y}

\input{tables/sp500_pval_2y}

\input{tables/sp500_pval_3y}

@@

\input{tables/sp500_pval_5y}

While we expect there to be differences in the p-values for the different 
methods, in this example, there are many instances where the conclusions
at the .05 significance level for our method are different than those
for both
\citep{babu2004goodness}'s and \citet{zeimbekakis2022misuses}'s method.
For instance, for a duration of 4 years, our method fails to reject that
the series follows the $v = 5$ Student's $t$ distribution with p-value .061,
wheraeas \citep{babu2004goodness}'s method finds a p-value of .0222, and 
\citet{zeimbekakis2022misuses}'s method finds a p-value of .0141. Although this
will not always be the case, we have found at least one example of a difference
in conclusions for different methods.

\subsection{SP100}
\label{sec:sp100}

Once again using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices for the S\&P 100 index. We varied the
durations as we did with the S\&P 500 index.
We approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$. 
We used $B = 10,000$ bootstrap 
replicates in this analysis as well.
The p-values for these tests are highlighted in 
Tables~\ref{table:SP1001}--\ref{table:SP1005}. One can observe more examples
of our method reaching different conclusions at the .05 level than others.
For a duration of 4 years, our method fails to reject that
the series follows the $v = 5$ Student's $t$ distribution with p-value .0873,
wheraeas \citep{babu2004goodness}'s method finds a p-value of .0368, and 
\citet{zeimbekakis2022misuses}'s method finds a p-value of .03. we have here
another instance where using a different method would lead to a different
conclusion about the marginal distribution of the population of interest.

\input{tables/sp100_pval_1y}

\input{tables/sp100_pval_2y}

\input{tables/sp100_pval_3y}

@@

\input{tables/sp100_pval_5y}

@@

The Kolmogorov-Smirnov test provides a way for researchers in a broad array
of applications to compare the distributions of populations. The
method proposed in this paper offers a variant of the test that can
be utilized for serially dependent data without specifying parameters.
Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. It requires a larger sample size 
than basic bootstrap to avoid either type I or type II errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust. We have also 
demonstrated possible applications of this method to financial data.
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap KS test and \citet{zeimbekakis2022misuses}'s parametric bootstrap
KS test, we have shown that our method can reach different conclusions which
would perhaps be missed if serial dependence was neglected and another
method was chosen.
This method can be used for any study where the goal is
to see if a serially dependent time series follows a hypothesized distribution
and the parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions and
different dependence structures. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

\jy{The number of references seems low. Need to cite more related works.}

\jy{Quality control the bib source}

>>>>

The Kolmogorov-Smirnov (KS) test is widely used to test if a sample is
from a given distribution. While a non-parametric bootstrap bias correction
exists for independent and identically distributed (i.i.d.) samples with unknown
parameters, a comparable correction suitable 
for serially dependent data is lacking.
This study proposes a bias correction
using non-parametric 
block bootstrap, designed to approximate the KS statistic in scenarios where
parameters are unspecified and data exhibit serial dependence. We provide
theoretical justification for this approach, then we evaluate its size
and power through simulation. Finally, we demonstrate applications of 
the method to
stock returns of the S\&P 100 and S\&P 500 indices.

@@

The standard one-sample Kolmogorov-Smirnov (KS) test is widely
recognized as an effective good-of-fit test for continuous distributions.
Consider $X_1, ..., X_n$, a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ that $X_i$'s follow a specific
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. This distribution function is
precisely computable using contemporary statistical software
\citep{marsaglia2003evaluating}. The versatility of the KS test allows its
application across various domains, such as analyzing cosmic microwave
background radiation \citep{naess2012application}, monitoring the count rate of
radioactive data \citep{aslam2020introducing}, gear condition monitoring
\citep{andrade2001gear}, studying images of breast cancer tumors
\citep{demidenko2004kolmogorov}, and examining financial markets
\citep{lux2001turbulence}.

Despite its widespread use, the KS test can be 
misapplied when its foundational assumptions are overlooked. The KS test 
assumes the data are independently and identically distributed (i.i.d.) and 
that the hypothesized distribution is continuous and fully specified without 
the need for parameter estimation. \citet{zeimbekakis2022misuses} examines 
common misapplications of the one-sample KS test, notably the inappropriate
use when the hypothesized distribution contains unspecified parameters.
For this situation, \citet{zeimbekakis2022misuses} proposed a parametric
bootstrap, where bootstrap samples of the test statistics are constructed from
samples generated from the fitted hypothesized distribution.
Alternatively, \citet{babu2004goodness} address this issue through the
application of basic non-parametric bootstrap methods.

Another prevalent misapplication of the KS test arises when data exhibit serial dependence, often overlooked in analyses. In an investigation of performance variation in
high-performance computing systems, \citet{tuncer2019ieee} did not describe
how or if they accounted for serial dependence when applying the two-sample
KS test.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.

A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

@@

The remainder of this paper is structured as follows: 
Section~\ref{sec:methods} provides an overview of the block bootstrap 
procedure and introduces the bias correction methodology for the 
non-parametric block bootstrap KS test. Section~\ref{sec:simu} is divided 
into two parts; initially, we evaluate the KS test's ability to maintain 
its size, i.e., its consistency in not rejecting the null hypothesis when 
it is indeed true. Subsequently, we examine the test's power, assessing 
whether it can reject the null hypothesis when it is false. Practical 
applications of our method are presented in Section~\ref{sec:real}, where 
we apply the approach to analyze if S\&P 100 and S\&P 500 index stock return data adhere to either the Normal 
or the Student's $t$ distribution. The 
paper concludes with Section~\ref{sec:conclusion}, offering final thoughts 
and remarks.

@@

First, let us review how the KS statistic is typically computed for a sample
with no serial dependence with fitted parameters.
Let $\hat\theta_n$ denote the fitted parameters for the hypothesized 
distribution fitted onto $X_t$, and let 
$F_n$ denote the empirical distribution function based on $X_1,...,X_n$.
Let
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x; \hat\theta_n)).
\end{equation*}
Then, the
goodness of fit statistic is 
\begin{equation*}
T_n := \sup_x|Y_n(x; \hat\theta)|.
\end{equation*}

@@

In this simulation study, our objective is twofold: firstly, to demonstrate that
under the null distribution, our test rejects the null hypothesis ($H_0$) 
approximately at the specified significance level. Secondly, we aim to
illustrate that under the alternate distribution, our test reliably rejects
$H_0$ at a high frequency. The fulfillment of both these criteria would indicate
the method's efficacy.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
\mc{addressed}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}$.

@@

Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices for the S\&P 500 index. We
observed how the method performs using the last 
4 years of data: January 1st, 2020 to December 31st, 2023.
We approximated stock returns by taking the difference in the logarithm 
of the closing prices.
Using our method, we tested that the stock returns are normally
distributed and Student's $t$ distributed 
with degrees of freedom $v \in \{30, 20, 10, 5, 4, 3, 2, 1\}$.
We tested the same hypotheses using \citet{babu2004goodness}'s 
non-parametric basic bootstrap bias correction and 
\citet{zeimbekakis2022misuses}'s parametric bootstrap method.
We used $B = 10,000$ bootstrap 
replicates in this analysis.
The p-values for these tests are highlighted in 
Table~\ref{table:SP5004}.

Once again, utilizing the \textsl{tseries} package \citep{tseries}, we collected
closing stock prices for the S\&P 100 index. Employing the same methodology as
our study on the S\&P 500 index, we conducted our analysis. The resulting 
p-values for these tests are presented in Table~\ref{table:SP1004}.

@@

@@

While differences in p-values among different methods are expected, our analysis
reveals numerous instances where conclusions at the 0.05 significance level
using our method differ from those obtained using the methods proposed by
\citet{babu2004goodness} and \citet{zeimbekakis2022misuses}.

For example, when assessing the S\&P 500 series, our method fails to reject the
hypothesis that the series follows a Student's $t$ distribution with degrees of
freedom $v = 5$ with a p-value of 0.061. In contrast, Babu and Sen's method
yields a p-value of 0.0222, and Zeimbekakis et al.'s method yields a p-value of
0.0141. Although such differences are not guaranteed, this
instance highlights a disagreement in conclusions among methods for the S\&P 500
index.

For the S\&P 100 index, our method demonstrates more instances where it reaches
different conclusions at the 0.05 significance level compared to other methods.
Our method fails to reject the hypothesis that the series conforms to a
Student's $t$ distribution with degrees of freedom $v = 5$, yielding a p-value
of 0.0873. In contrast, \citet{babu2004goodness}'s method yields a p-value of 
0.0368, and \citet{zeimbekakis2022misuses}'s method yields a p-value of 0.03. 
We have here
another instance where employing different methods leads to divergent
conclusions regarding the marginal distribution of the population under scrutin.

@@

The Kolmogorov-Smirnov test stands as a crucial tool for researchers across
diverse fields, enabling comparisons of population distributions. This paper
introduces a variant of the test tailored for serially dependent data without
necessitating parameter specification.
Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. Notably, it demands a larger sample
size compared to basic bootstrap methods to avoid both type I and type II 
errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust.

We have also 
demonstrated possible applications of this method to financial data.
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap bias correction and \citet{zeimbekakis2022misuses}'s parametric
bootstrap method, we have underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions and
different dependence structures. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

\jy{Quality control the bib source}
\mc{addressed}
